Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-10-01 Rebuttal/Rebuttal.tex

\textbf{Answer:}

We thank the reviewer for raising this interesting point of discussion. The scalability of the problem when dealing with more complex vehicle models is a challenging aspect of the procedure: in fact, considering a $n$-states vehicle model, the resulting stochastic OCP system has $n+\frac{n(n+1)}{2}$ states.

The simplified model used in the current implementation fits well to the serial approach of the OCP, since computational times and total number of decision variables are still sufficiently small.

In further works we are actually planning to adopt more complex models and to parallelize the problem solution on multiple cores. This approach has been explored by our research group in these previous works:

\begin{itemize}
  \item https://doi.org/10.1007/s11044-023-09932-6
  \item https://doi.org/10.1080/00423114.2024.2407175,
\end{itemize}

where the nominal minimum lap-time problem was solved for a 14-DoF multibody model.

Finally, since the reviewer has pointed out this specific feature, we would like to remark that, although the out-of-plane dynamics is neglected in our model, we took at least into account the (steady-state) longitudinal load transfers.

@@

\item
\textit{Sec. 2.4, page 7: as shown by the authors, considering the robustness of the friction limit is relatively straightforward when using the friction-ellipse constraint. How can this aspect be addressed when tyre coupling is implicitly captured by the magic formula (i.e. when no explicit friction-ellipse constraint is imposed)?}

\vspace{2mm}

\textbf{Answer:}

We sincerely thank the reviewer for the stimulating question. When tyre coupling is modeled implicitly via the Magic Formula (MF) in combined slip, the friction boundary is not an explicit ellipse but a nonlinear frontier emerging from
\begin{equation}
F_x = G_{xa}(\alpha,\kappa,\ldots)\,F_{x0}(\kappa),
\;
F_y = G_{yk}(\alpha,\kappa,\ldots)\,F_{y0}(\alpha) + SV_{yk}(\alpha,\kappa,\ldots),
\label{eq:MF-combined}
\end{equation}
where $F_{x0},F_{y0}$ are the pure‐slip curves (with peak factors $D_x,D_y$ depending on $F_z,\gamma,\ldots$), $G_{xa},G_{yk}$ are the normalized “hill–shaped” coupling weights (equal to $1$ in pure slip), and $SV_{yk}$ is the shift induced by ply–steer/conicity effects/rolling resistance shift, see e.g. [Pacejka 2012, 3rd edition, p.166].
To retain the same \emph{robustness to friction variations} that is straightforward with an explicit friction ellipse, we could introduce a \emph{control-level} constraint built from the instantaneous MF capacities (a sort of \emph{breathing ellipse}).

\vspace{4pt} 
Given force references $(F_x^{\mathrm{ref}},F_y^{\mathrm{ref}})$ produced at the generic iteration of the NLP solver, we could define the instantaneous directional limits from the MF model at the current $(\alpha,\kappa)$:
\begin{equation}
F_{x,\lim} \;=\; D_x\,G_{xa}(\alpha,\kappa,\ldots),
\qquad
F_{y,\lim} \;=\; D_y\,G_{yk}(\alpha,\kappa,\ldots).
\label{eq:limits}
\end{equation}
Of course, since $SV_{yk}$ is an offset, we could work with the lateral \emph{effective} command $F_{y,\mathrm{eff}}^{\mathrm{ref}} := F_y^{\mathrm{ref}} - SV_{yk}$.
Then the new \emph{axle saturation ratio} $S$ would read
\begin{equation}
S=
\left(\frac{F_x^{\mathrm{ref}}}{F_{x,\lim}}\right)^2
+
\left(\frac{F_{y,\mathrm{eff}}^{\mathrm{ref}}}{F_{y,\lim}}\right)^2,
\label{eq:util}
\end{equation}
and robustness to friction uncertainty would be enforced by the same considerations that have been proposed in Sec. 2.4 of the paper.
It is worth observing that this \emph{breathing ellipse} would not add additional physics: the MF already encodes tyre coupling. It would just provide a control–level robustness margin in analogy to the classical friction‐ellipse constraint, but \emph{coherent} with the instantaneous MF capacities $D\,G$ (and with $SV_{yk}$ excluded from the lateral capacity).

@@

\item
\textit{Sec. 2.5, page 8: eq. (11) introduces an algebraic variable e, representing the lateral position of the vehicle with respect to the centreline. Why not use e as a vehicle state instead of x, y, thereby avoiding algebraic varibles?}

\vspace{2mm}

\textbf{Answer:}

We thank the reviewer for the insightful question. The main motivation for our current formulation is that the vehicle dynamics are not embedded in the track frame. In earlier work we did adopt a track-embedded type formulation (see A reduced-order Lie-group–based race car model for systematic trajectory optimization on 3D tracks, Meccanica 58:1869–1883, 2023, doi:10.1007/s11012-023-01708-8.) where $e$ ($q_2$ in the mentioned paper) would emerge naturally as a parametrization coordinate.

Within a direct-collocation + IPOPT pipeline, keeping $(x,y)$ as states and also adding $e$ as algebraic variable, yields Jacobians that are sparser and typically less involved algebraically: the track geometry (centerline curvature/banking and their derivatives) enters only the algebraic relation for $e$, rather than propagating through the full state dynamics at every collocation point. By contrast, promoting $e$ to a state in a track-embedded (Frenet-Serret type) model couples the dynamics to space-varying geometry terms, which tends to densify the KKT system and degrade conditioning. A further practical advantage is that many constraints and features (e.g., global track limits, obstacles, multi-vehicle interactions) are naturally expressed in inertial coordinates and remain simpler when $(x,y)$ are states.

@@

\item
\textit{Sec. 4.4, page 13: eq (20) introduces smoothed approaximation functions to ensure differentiatibility, which depend on a sharpness parameter $\chi$. Which are the typical values used in the authours' implementation?}

\vspace{2mm}

\textbf{Answer:}

We appreciate the reviewer's request \ul{for more details on these technical aspects.} In this context, we chose a value of $\chi=10$ as the sharpness parameter, and we have explicitly stated it in the paper. This value proved to be a good trade-off between a steep transition and the smoothness required by the solver to enhance convergence. In general, $\chi$ must be tuned according to the range of the corresponding variable. In this case, the longitudinal force $F_x$ is on the order of $10^4$\,N.

### 2025-10-01 performance_comparison.tex
[mg] Paper edits Rev2-Ans1,3,6,7 + bold P pag 8 + old comments cleaning

First we present selected results from the open-loop planning, which helps illustrate the method and its underlying mechanism.
%We begin by presenting a selection of results obtained with the open-loop planning approach, which allows for a clearer illustration of the method and its underlying mechanisms.
This serves as a reference to assess the impact of key design parameters and constraint types on planned trajectories.
We then compare open- and closed-loop strategies, and finally validate the framework by comparing planned trajectories and simulations with noise realization.
%A comparison is then carried out between the open-loop and closed-loop strategies. Finally, the planned trajectories are compared against those followed in simulations with noise realizations, providing empirical validation of the proposed framework.

@@

% Contour plot dei LAP time in funzione dei due gamma, lo mettiamo o no? Eventialmente nel comparison open/closed

@@

\subsection{Open-loop parameter sensitivity}
\label{sec:ol_param_sensitivity}
To better understand the effect of key parameters, we focus on the open-loop formulation with robustified track limit constraint. This choice favors clearer visual interpretation.
%To gain deeper insight into how key parameters affect the planning outcome, we focus here on the open-loop formulation. At this stage, only the track limit constraint is robustified. This choice is motivated by the fact that its effects are more visually evident than those of the friction limit constraint, thereby facilitating a more immediate understanding of their impact.
The first parameter considered is the confidence level of constraint satisfaction.
We recall that the factor $\ga^\textrm{TLC}$ acts as a tuning knob, multiplying the standard deviation of the constraint, $\sigma^\textrm{TLC}$, to determine the total back-off term $\be^\textrm{TLC}$.
This parameter directly influences the probability of satisfying the track limit constraint: higher values of $\ga^\textrm{TLC}$ correspond to more conservative (i.e., robust) behavior.
Specifically, $\ga^\textrm{TLC} = 0$ yields a satisfaction probability of 50\% and leads to the nominal solution with no back-off, while $\ga^\textrm{TLC} = 3$ corresponds to a confidence level of 99\%.
%For the purposes of this analysis, we examine the effect of varying $\ga^\textrm{TLC}$ in this range.

%Setting $\ga^\textrm{TLC} = 0$ leads to the nominal solution. In this case, the back-off term $\be^\textrm{TLC}$ becomes zero, meaning that the constraint is not robustified and thus coincides with the original, non-robust formulation.

@@

\begin{figure}
	\centering
	\includegraphics{Fig/gamma_sensitivity.pdf}
	\hfill
	\includegraphics{Fig/Pzero_sensitivity.pdf}
\caption{Effect of the TLC penalty coefficient $\gamma^{\mathrm{TLC}}$ (left panel) and the initial covariance $\mathbf P_{0}$ (right panel) on the optimal trajectory.
Raising $\gamma^{\mathrm{TLC}}$ from 0 to 3 increases the probability of satisfying the track-limit constraints from 50 \% to 99 \%.
The right panel compares the nominal covariance $\bar{\mathbf P}_{0}$ (black) with $4\bar{\mathbf P}_{0}$ (blue line) and $\tfrac14\,\bar{\mathbf P}_{0}$ (red line), i.e., twice and half the nominal state uncertainty.
}
%\caption{Effect of $\ga^\textrm{TLC}$ (left panel) and $\bP_0$ (right panel) on optimal trajectory. The parameter $\ga^\textrm{TLC}$ spans in the range $\left[0,3\right]$ corresponding to a probability to meet the track limit constraint ranging from 50\% to 99\%. In the right panel, in addition to the baseline value $\bar{\bP}_0$ (black line), two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ (blue line) and $\bP_0 = \frac{1}{4}\bar{\bP}_0$ (red line). These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector.}
	\label{fig:ol_sensitivities}
\end{figure}

@@

 Assuming the state vector is ordered as described in Sec.~\ref{sec:vehicle_model}, the baseline initial covariance matrix is set to $\bP_0=\diag(\bbsig^2)$, with
\begin{align}
\bar{\boldsymbol{\sigma}}~=~\left[0.1\,\textrm{m/s}, 0.01\,\textrm{m/s}, 0.01\,\textrm{rad/s}, 1\,\textrm{m}, 1\,\textrm{m}, 0.0175\,\textrm{rad}\right]^T.
\end{align}
These standard deviations have been chosen based on the typical range of variation of each state. To provide a concrete interpretation at the starting condition, the assumption on first state $u \sim \calN(\mu_u,\sig_u^2)$, with $\sig_u=0.1$ m/s, implies that
$\Pr\{\mu_u-\sig_u \leq u \leq \mu_u+\sig_u\}\approx 68\%$ and $\Pr\{\mu_u-3\sig_u \leq u \leq \mu_u+3\sig_u\}\approx 99\%$.

@@

	%\caption{Evolution of the longitudinal speed $\De u$, wheel steering angle $\De \de$, longitudinal force $\De F_x$, and lateral deviation $\De e$ are reported for three robustification scenarios: TLC (Track Limit Constraints), FLC (Friction Limit Constraint), and TLC+FLC (both constraints enforced concurrently). Each profile is shown relative to the baseline (nominal) trajectory to highlight the effect of each constraint-handling strategy. The dashed lines in the last three panels indicate the sign of the reference signal, allowing one to determine whether a positive variation with respect to the baseline corresponds to an increase or a decrease in the absolute value. This distinction is particularly important for quantities that depend on the direction of the turn, such as the wheel steering angle $\de$ and the lateral deviation $e$, while in the case of $F_x$, it indicates whether the vehicle is accelerating or braking.}
	\label{fig:ol_telemetries}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[scale = 1]{Fig/ol_saturation.pdf}
	\caption{Open-Loop. Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear parameter $\al$. Nominal configuration in dashed black, TLC in green, FLC in red, TLC+FLC in blue.
	%The dashed black line corresponds to the reference (nominal) configuration, while the green, red, and blue lines represent the robustified configurations TLC, FLC, and TLC+FLC, respectively.
	}
	\label{fig:ol_saturation}
\end{figure}

%It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
We now investigate how the inclusion of different kinds of robustified constraints --- individually or in combination, as described in Sec.~\ref{sec:FLC} and~\ref{sec:TLC} --- influences the driving style.
We compare the nominal feed-forward optimal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC).

First we show results obtained with the open-loop method described in Sec.~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
%All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
%---Delta key signals
The comparison between the nominal case and the three robustified cases is illustrated in Fig.~\ref{fig:ol_telemetries}, in terms of the deviation of selected signals from to their nominal counterparts.
The panels display variations on the longitudinal speed $\De u$ (first panel), wheel steering angle $\De \de$ (second panel), total longitudinal force $\De F_x$ (third panel), and lateral deviation of the center of mass (CoM) from the track centerline $\De e$ (fourth panel).
Except for the $\De u$ panel, each plot includes dashed black lines that indicate the sign of the nominal signal.
These indicators are shown at three distinct vertical levels --- high, medium, and low --- corresponding respectively to positive, near-zero, and negative values of the nominal signal.

@@

This shift in trajectory and longitudinal speed becomes especially evident when approaching the sharp turn for $\al \in [0.707, 0.712]$. A noticeable increase in longitudinal speed is observed around $\alpha = 0.71$ for the TLC configuration  in the $\De u$ panel. This peak is followed by a sharp drop in speed, indicating intense braking.
The TLC+FLC configuration demonstrates a qualitatively similar trend in $\Delta u$, although the peak is significantly lower and the subsequent speed reduction is clearly stronger. This behavior is consistent with the variation of the total longitudinal force $\Delta F_x$ shown in the third panel.
%\textcolor{red}{In particular, the TLC and TLC+FLC configurations exhibit a higher longitudinal force while approaching the sharp turn, in the curvilinear abscissa interval $\al\in\left[0.707, 0.712\right]$, resulting in a higher longitudinal speed.
%Subsequently, these two configurations apply a lower longitudinal force during braking compared to the nominal case, indicating more intense braking. This behavior explains the sharp reduction in longitudinal speed observed thereafter.}

@@

%As already observed, the configurations with the robustified track limit constraint (green and blue lines) follow significantly different paths compared to the nominal solution. This deviation results in distinct ground force demands, clearly visible in the saturation index.
%From this plot, it can be observed that the configurations with the robustified track limit constraint (green and blue lines) follow significantly different trajectories compared to the nominal solution, resulting in distinct ground force demands. This behavior is noticeable in turn 1 and even more clearly in turn 2. Specifically, these two configurations begin to experience high axle loads approximately 12\,m before the nominal trajectory and sustain them for an additional 4.6\,m (TLC, green line) and 9.2\,m (TLC+FLC, blue line) beyond the nominal reference.

@@

	%\caption{Evolution of the longitudinal speed $\De u$, wheel steering angle $\De \de$, longitudinal force $\De F_x$, and lateral deviation $\De e$ are reported for three robustification scenarios: TLC (Track Limit Constraints), FLC (Friction Limit Constraint), and TLC+FLC (both constraints enforced concurrently). Each profile is shown relative to the baseline (nominal) trajectory to highlight the effect of each constraint-handling strategy. The dashed lines in the last three panels indicate the sign of the reference signal, allowing one to determine whether a positive variation with respect to the baseline corresponds to an increase or a decrease in the absolute value. This distinction is particularly important for quantities that depend on the direction of the turn, such as the wheel steering angle $\de$ and the lateral deviation $e$, while in the case of $F_x$, it indicates whether the vehicle is accelerating or braking.}
	\label{fig:cl_telemetries}
\end{figure}
\begin{figure}[b!]
	\centering
	\includegraphics[scale = 1.005]{Fig/cl_saturation.pdf}
	\caption{Closed-Loop. Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear parameter $\al$. Nominal configuration in dashed black, TLC in green, FLC in red, TLC+FLC in blue.
		%The dashed black line corresponds to the reference (nominal) configuration, while the green, red, and blue lines represent the robustified configurations TLC, FLC, and TLC+FLC, respectively.
	}
	\label{fig:cl_saturation}
\end{figure}

@@

%\subsection{Open-loop vs closed-loop robust planning}
\begin{figure}[b!]
	\centering
	\includegraphics{Fig/olcl_traj_bands.pdf}
	\caption{Back-off envelopes for the open-loop (red) and closed-loop (cyan) robust planners. Each shaded band represents the area swept by the back-off-extended vehicle and must lie entirely within the actual track boundaries.
		The open-loop approach yields a wider envelope due to worst-case growth assumptions, whereas the closed-loop design, aided by time-varying LQR controller, maintains a narrower margin.}
	\label{fig:olcl_traj_bands}
\end{figure}

@@

\subsection{Simulations with noise realization}
\begin{figure}[t!]
	\centering
	\includegraphics{Fig/olcl_traj_strings.pdf}
	\caption{100 simulations (each panel) with random initial conditions and Gaussian noise realizations. In the left panel, the reference trajectory (black curve) is the non-robust one, tracked using an LQR controller designed around it as per Step 2 of Subsection~\ref{sec:LQR}. In the right panel, the reference (black curve) is a robust trajectory obtained using the closed-loop optimization method as per Step 3 of Subsection~\ref{sec:CLrobustified}, which is tracked using the corresponding optimized controller.}
	\label{fig:traj_strings}
\end{figure}
To provide an empirical validation of the closed-loop robustified planning we simulated the evolution of the controlled vehicle with 100 random initial conditions and included noise realizations. The random values for the initial state and the noise follow a Gaussian distribution with covariance matrices $\bP_0$ and $\bQ$, in both cases, as introduced in Eq.~\eqref{eq:dP}. This analysis is not applicable the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon.
%For visual clarity, the comparison is limited to trajectories with the robustified TLC, but analogous considerations apply to the robustified FLC as well.

@@

%---------------------------------
%ORIGINAL VERSION

%This analysis aims to provide empirical validation of the robust trajectories obtained using the methods proposed in this work. For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
%An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented to stabilize the nominal trajectory. In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.

>>>>

First we present selected results from the open-loop planning, which helps illustrate the method and its underlying mechanism.
This serves as a reference to assess the impact of key design parameters and constraint types on planned trajectories.
We then compare open- and closed-loop strategies, and finally validate the framework by comparing planned trajectories and simulations with noise realization.

@@

@@

\subsection{Open-loop parameter sensitivity}
\label{sec:ol_param_sensitivity}
To better understand the effect of key parameters, we focus on the open-loop formulation with robustified track limit constraint. This choice favors clearer visual interpretation.
The first parameter considered is the confidence level of constraint satisfaction.
We recall that the factor $\ga^\textrm{TLC}$ acts as a tuning knob, multiplying the standard deviation of the constraint, $\sigma^\textrm{TLC}$, to determine the total back-off term $\be^\textrm{TLC}$.
This parameter directly influences the probability of satisfying the track limit constraint: higher values of $\ga^\textrm{TLC}$ correspond to more conservative (i.e., robust) behavior.
Specifically, $\ga^\textrm{TLC} = 0$ yields a satisfaction probability of 50\% and leads to the nominal solution with no back-off, while $\ga^\textrm{TLC} = 3$ corresponds to a confidence level of 99\%.

@@

\begin{figure}
	\centering
	\includegraphics{Fig/gamma_sensitivity.pdf}
	\hfill
	\includegraphics{Fig/Pzero_sensitivity.pdf}
\caption{Effect of the TLC penalty coefficient $\gamma^{\mathrm{TLC}}$ (left panel) and the initial covariance $\mathbf P_{0}$ (right panel) on the optimal trajectory.
Raising $\gamma^{\mathrm{TLC}}$ from 0 to 3 increases the probability of satisfying the track-limit constraints from 50 \% to 99 \%.
The right panel compares the nominal covariance $\bar{\mathbf P}_{0}$ (black) with $4\bar{\mathbf P}_{0}$ (blue line) and $\tfrac14\,\bar{\mathbf P}_{0}$ (red line), i.e., twice and half the nominal state uncertainty.
}

	\label{fig:ol_sensitivities}
\end{figure}

@@

 Assuming the state vector is ordered as described in Sec.~\ref{sec:vehicle_model}, the baseline initial covariance matrix is set to $\bP_0=\diag(\bbsig)^2$, with
\begin{align}
\bar{\boldsymbol{\sigma}}~=~\left[0.1\,\textrm{m/s}, 0.01\,\textrm{m/s}, 0.01\,\textrm{rad/s}, 1\,\textrm{m}, 1\,\textrm{m}, 0.0175\,\textrm{rad}\right]^T.
\end{align}
These standard deviations have been chosen based on the typical range of variation of each state. To provide a concrete interpretation at the starting condition, the assumption on first state $u \sim \calN(\mu_u,\sig_u^2)$, with $\sig_u=0.1$ m/s, implies that
$\Pr\{\mu_u-\sig_u \leq u \leq \mu_u+\sig_u\}\approx 68\%$ and $\Pr\{\mu_u-3\sig_u \leq u \leq \mu_u+3\sig_u\}\approx 99\%$.

@@

	\label{fig:ol_telemetries}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[scale = 1]{Fig/ol_saturation.pdf}
	\caption{Open-Loop. Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear parameter $\al$. Nominal configuration in dashed black, TLC in green, FLC in red, TLC+FLC in blue.
	}
	\label{fig:ol_saturation}
\end{figure}

We now investigate how the inclusion of different kinds of robustified constraints --- individually or in combination, as described in Sec.~\ref{sec:FLC} and~\ref{sec:TLC} --- influences the driving style.
We compare the nominal feed-forward optimal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC).

First we show results obtained with the open-loop method described in Sec.~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.

%---Delta key signals
The comparison between the nominal case and the three robustified cases is illustrated in Fig.~\ref{fig:ol_telemetries}, in terms of the deviation of selected signals from to their nominal counterparts.
The panels display variations on the longitudinal speed $\De u$ (first panel), wheel steering angle $\De \de$ (second panel), total longitudinal force $\De F_x$ (third panel), and lateral deviation of the center of mass (CoM) from the track centerline $\De e$ (fourth panel).
Except for the $\De u$ panel, each plot includes dashed black lines that indicate the sign of the nominal signal.
These indicators are shown at three distinct vertical levels --- high, medium, and low --- corresponding respectively to positive, near-zero, and negative values of the nominal signal.

@@

This shift in trajectory and longitudinal speed becomes especially evident when approaching the sharp turn for $\al \in [0.707, 0.712]$. A noticeable increase in longitudinal speed is observed around $\alpha = 0.71$ for the TLC configuration  in the $\De u$ panel. This peak is followed by a sharp drop in speed, indicating intense braking.
The TLC+FLC configuration demonstrates a qualitatively similar trend in $\Delta u$, although the peak is significantly lower and the subsequent speed reduction is clearly stronger. This behavior is consistent with the variation of the total longitudinal force $\Delta F_x$ shown in the third panel.

@@

@@

	\label{fig:cl_telemetries}
\end{figure}
\begin{figure}[b!]
	\centering
	\includegraphics[scale = 1.005]{Fig/cl_saturation.pdf}
	\caption{Closed-Loop. Axle saturation $S_j(\bx,\bu)$ for front axle (j=1, first panel) and rear axle (j=2, second panel) with respect to the curvilinear parameter $\al$. Nominal configuration in dashed black, TLC in green, FLC in red, TLC+FLC in blue.
	}
	\label{fig:cl_saturation}
\end{figure}

@@

\begin{figure}[b!]
	\centering
	\includegraphics{Fig/olcl_traj_bands.pdf}
	\caption{Back-off envelopes for the open-loop (red) and closed-loop (cyan) robust planners. Each shaded band represents the area swept by the back-off-extended vehicle and must lie entirely within the actual track boundaries.
		The open-loop approach yields a wider envelope due to worst-case growth assumptions, whereas the closed-loop design, aided by time-varying LQR controller, maintains a narrower margin.}
	\label{fig:olcl_traj_bands}
\end{figure}

@@

\subsection{Simulations with noise realization}
\begin{figure}[t!]
	\centering
	\includegraphics{Fig/olcl_traj_strings.pdf}
	\caption{100 simulations (each panel) with random initial conditions and Gaussian noise realizations. In the left panel, the reference trajectory (black curve) is the non-robust one, tracked using an LQR controller designed around it as per Step 2 of Subsection~\ref{sec:LQR}. In the right panel, the reference (black curve) is a robust trajectory obtained using the closed-loop optimization method as per Step 3 of Subsection~\ref{sec:CLrobustified}, which is tracked using the corresponding optimized controller.}
	\label{fig:traj_strings}
\end{figure}
To provide an empirical validation of the closed-loop robustified planning we simulated the evolution of the controlled vehicle with 100 random initial conditions and included noise realizations. The random values for the initial state and the noise follow a Gaussian distribution with covariance matrices $\bP_0$ and $\bQ$, in both cases, as introduced in Eq.~\eqref{eq:dP}. This analysis is not applicable the open-loop approach, as it does not compute a stabilizing controller---resulting in trajectories that would significantly deviate from the track limits beyond the prediction horizon.

@@
